Model Search in Relational Domains

### 2016-10-22 example.tex
right align numbers
nudge Fig 1

\section{Motivating Example}
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & Control-Flow Graph & Abstract Conflict Graph \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void foo(int v, 
         _Bool c) 
{
  assume(v>=0 && v<=N);
  int x; 
  if(c)
    x = v;
  else 
    x = -v;
  int z = x * x;
  assert(z>=0);
}
\end{lstlisting}
&
\begin{minipage}{3.7cm}
\centering
%\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{example.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
&
\begin{minipage}{4.50cm}
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{acdl_run.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
\\
\hline
\end{tabular}
\caption{\label{fig:example}
C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
\end{figure}
Let us consider a simple program in Figure~\ref{fig:example} that performs 
multiplication of two machine integers and checks that the result of the 
multiplication is always positive for an integer input $v$.  To avoid 
overflow error, we assume $v$ has an upper bound $N$.  The control 
flow graph is shown in the middle column of Figure~\ref{fig:example}.

We analyse the program using three different analysers -- {\em bounded 
model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
\Omit {
The SAT solver makes 36 decisions, 1365 Boolean constraint propagations 
and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
the solver performs a restart, presumably to recover from bad decisions.  
}
Standard BMC tools convert the program to a bit-vector equation and pass 
them to a SAT solver.  However, standard forward interval analysis is 
too imprecise to verify the safety of the above program.  The 
imprecision is due to control-flow join at node $n4$.  State-of-the-art 
abstract interpretation tool, Astr{\'e}e \cite{se2011}, requires external hints, provided 
by manually annotating the code with partition directives at $n1$, 
to prove safety.  The partition directive helps the tool to analyse 
each program execution paths separately.  To do so, Astr{\'e}e 
requires a product of trace partitioning domain and an interval domain 
to determine safety.  However, a relational domain such as octagons is 
able to prove the program without any partition.  
\Omit {
In general, the imprecision is either intended by the tool because such high 
precision analysis is normally not required for runtime error 
analysis or the imprecision is unavoidable due to the complexity 
of the application under analysis.  
}
%
\begin{table}[!b]
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline
ACDCL (Octagons) & 0 & 7 & 0 & 0 & 0 \\ 
\hline
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{solver}
\end{table}
%
The right-hand side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
octagon domain (top) and an interval domain (bottom).  The analysis 
associates an abstract element for each control location and variable, 
as shown in the abstract conflict graph in Figure~\ref{fig:example}.  
Octagonal constraints deduced by ACDCL solver are sufficient to prove 
safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
an automatic interval analysis is also sufficient to prove safety with 
only a single decision.  Interval constraints generated from forward 
analysis in the initial deduction phase infers that 
$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL cannot prove safety.

>>>>

\section{Motivating Example}

\begin{figure}[t]
\scriptsize\centering
\begin{tabular}{c|c|c}
\hline
C program & Control-Flow Graph & Abstract Conflict Graph \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void foo(int v, 
         _Bool c) 
{
  assume(v>=0 &&
         v<=N);
  int x; 
  if(c)
    x = v;
  else 
    x = -v;
  int z = x * x;
  assert(z>=0);
}
\end{lstlisting}
&
\begin{minipage}{3.7cm}
\centering
%\vspace*{0.3cm}
\scalebox{.52}{\import{figures/}{example.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
&
\begin{minipage}{5.6cm}
\centering
\vspace*{0.3cm}
\scalebox{.5}{\import{figures/}{acdl_run.pspdftex}}
%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{minipage}
\\
\hline
\end{tabular}
\caption{\label{fig:example}
C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
\end{figure}

Let us consider a simple program in Figure~\ref{fig:example} that performs 
multiplication of two machine integers and checks that the result of the 
multiplication is always positive for an integer input $v$.  To avoid 
overflow error, we assume $v$ has an upper bound $N$.  The control 
flow graph is shown in the middle column of Figure~\ref{fig:example}.

We analyse the program using three different analysers -- {\em bounded 
model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
\Omit {
The SAT solver makes 36 decisions, 1365 Boolean constraint propagations 
and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
the solver performs a restart, presumably to recover from bad decisions.  
}
Standard BMC tools convert the program to a bit-vector equation and pass 
them to a SAT solver.  However, standard forward interval analysis is 
too imprecise to verify the safety of the above program.  The 
imprecision is due to control-flow join at node $n4$.  State-of-the-art 
abstract interpretation tool, Astr{\'e}e \cite{se2011}, requires external hints, provided 
by manually annotating the code with partition directives at $n1$, 
to prove safety.  The partition directive helps the tool to analyse 
each program execution paths separately.  To do so, Astr{\'e}e 
requires a product of trace partitioning domain and an interval domain 
to determine safety.  However, a relational domain such as octagons is 
able to prove the program without any partition.  
\Omit {
In general, the imprecision is either intended by the tool because such high 
precision analysis is normally not required for runtime error 
analysis or the imprecision is unavoidable due to the complexity 
of the application under analysis.  
}
%
\begin{table}[!b]
\begin{center}
{
\begin{tabular}{l|r|r|r|r|r}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline
ACDCL (Octagons) & 0 & 7 & 0 & 0 & 0 \\ 
\hline
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{solver}
\end{table}
%
The right-hand side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
octagon domain (top) and an interval domain (bottom).  The analysis 
associates an abstract element for each control location and variable, 
as shown in the abstract conflict graph in Figure~\ref{fig:example}.  
Octagonal constraints deduced by ACDCL solver are sufficient to prove 
safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
an automatic interval analysis is also sufficient to prove safety with 
only a single decision.  Interval constraints generated from forward 
analysis in the initial deduction phase infers that 
$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL cannot prove safety.
